---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rfc86

<!-- badges: start -->
<!-- badges: end -->

The goal of rfc86 is to add some benchmarks for reading files via the [proposed GDAL columnar API](https://github.com/rouault/gdal/blob/rfc_86/doc/source/development/rfc/rfc86_column_oriented_api.rst).

## Installation

You can install the development version of rfc86 from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("paleolimbot/rfc86")
```

To use it you will have to build GDAL from source against the [prototype implementation](https://github.com/rouault/gdal/tree/arrow_batch_new).

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(rfc86)
library(geoarrow)
library(sf)

if (!file.exists("nshn_water_line.parquet")) {
  curl::curl_download(
    "https://github.com/paleolimbot/geoarrow-data/releases/download/v0.0.1/nshn_water_line.parquet",
    "nshn_water_line.parquet"
  )
  
  curl::curl_download(
    "https://github.com/paleolimbot/geoarrow-data/releases/download/v0.0.1/nshn_water_line.gpkg",
    "nshn_water_line.gpkg"
  )
  
  system("ogr2ogr nshn_water_line.fgb nshn_water_line.gpkg")
}

system.time(read_ogr_table("nshn_water_line.gpkg"))
system.time(read_ogr_table("nshn_water_line.fgb"))
system.time(read_ogr_sf("nshn_water_line.gpkg"))
system.time(read_sf("nshn_water_line.gpkg"))

system.time(read_ogr_table("nshn_water_line.parquet"))
system.time(arrow::read_parquet("nshn_water_line.parquet"))
```

```{r}
# validate the output
table <- read_ogr_table("nshn_water_line.fgb")
table$ValidateFull()

table <- read_ogr_table("nshn_water_line.gpkg")
table$ValidateFull()

table <- read_ogr_table("nshn_water_line.parquet")
table$ValidateFull()
```

Files from https://github.com/paleolimbot/geoarrow-data/

```{r, eval=FALSE}
files <- list.files("../geoarrow-public-data/release-files", full.names = TRUE)
for (f in files) {
  message(sprintf("Checking '%s'...", basename(f)), appendLF = FALSE)
  table <- read_ogr_table(f)
  message(sprintf("%s...", table$ValidateFull()), appendLF = FALSE)
  
  tf <- tempfile(fileext = ".feather")
  arrow::write_feather(table, tf)
  table <- read_ogr_table(tf)
  message(sprintf("feather: %s", table$ValidateFull()))
  unlink(tf)
}
```

(All files pass individually but I can get a crash if I try to read them all sequentially)
